---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# lergm: Estimation of Little 'ERGMs' using exact likelihood

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) 
[![CRAN status](https://www.r-pkg.org/badges/version/lergm)](https://cran.r-project.org/package=lergm)
[![Travis build status](https://travis-ci.org/USCCANA/lergm.svg?branch=master)](https://travis-ci.org/USCCANA/lergm)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/USCCANA/lergm?branch=master&svg=true)](https://ci.appveyor.com/project/USCCANA/lergm)
[![Coverage status](https://codecov.io/gh/USCCANA/lergm/branch/master/graph/badge.svg)](https://codecov.io/github/USCCANA/lergm?branch=master)



The development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("USCCANA/lergm")
```
## Example

An example from the manual

## When `ergm` is not enough

```{r net1}
library(lergm)
library(sna)

# Generating a small graph
set.seed(12)
n <- 4
net <- sna::rgraph(n, tprob = .7)
gplot(net)
```

```{r comparing-w-ergm}
model <- net ~ edges + mutual + balance

library(ergm)
ans_lergm <- lergm(model)
ans_ergm  <- ergm(model)

# The lergm should have a larger value
ergm.exact(ans_lergm$coef, model)
ergm.exact(ans_ergm$coef, model)

summary(ans_lergm)
summary(ans_ergm)
```

Checking convergence diagnostics

```{r convergence-diag}
plot(ans_lergm)
```


## Do we get the same?

```{r net2}
# Generating a small graph
set.seed(12123)
n   <- 4
net <- sna::rgraph(n, tprob = .3)
gplot(net)
```

```{r}
model <- net ~ edges + mutual

library(ergm)
ans_lergm <- lergm(model)
ans_ergm  <- ergm(model, control = control.ergm(
  MCMC.effectiveSize = 4000,
  seed = 444)
  )

# The lergm should have a larger value
ergm.exact(ans_lergm$coef, model) > ergm.exact(ans_ergm$coef, model)

summary(ans_lergm)
summary(ans_ergm)
```

# Similarity indices

https://cran.r-project.org/web/packages/proxy/proxy.pdf

A Survey of Binary Similarity and Distance Measures
Seung-Seok Choi, Sung-Hyuk Cha, Charles C. Tappert
Department of Computer Science, Pace University
New York, US

# Contributing

Please note that the 'lergm' project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.