---
title: "CSS Analysis"
author: "George G. Vega Yon"
date: "October 22, 2018"
output:
  pdf_document:
    keep_tex: true
    md_extensions: "+autolink_bare_uris+ascii_identifiers+tex_math_single_backslash+raw_tex"
header-includes:
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
---

<!-- Default sans serif -->
\renewcommand{\familydefault}{\sfdefault} 

# Locally Aggregated Structures (LAS)

*   For each group $g \in G$ we generated what the literature calls Locally Aggregated Structure networks (LAS networks).

*   A tie $(i,j)$ in the LAS exists if and only if $(i,j) \in CSS_i$ and $(i,j) \in CSS_j$, i.e. if both $i$ and $j$ report the existance of such tie.

# Levels of agreement

*   For each $i \in N_g$ (individual $i$ in group $G$), we calculated the following metrics

```{r setup, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message = FALSE, autodep = -1L)
library(dplyr)
library(tidyselect)
library(tidyr)
library(magrittr)
library(similR)
library(igraph)
statistics <- c(
  `Hamman (S)`  = "shamann",
  `Hamming (D)` = "dhamming",
  `Mean Manhattan (D)` = "dmh",
  `Michael (S)` = "smichael",
  `Sized Difference (D)` = "dsd"
  )
```

```{r data, echo=FALSE}
# Reading covariate data
dat_group <- haven::read_spss("../data-raw/MURI_AllSurveys - FINAL - Group level data_1.sav")
dat_individual <- 
  haven::read_spss("../data-raw/MURI_AllSurveys - FINAL_073018.sav")
dat_individual <- dat_individual %>%
  mutate(Group = as.integer(Group))
```

```{r network-data, echo=FALSE}
# Reading raw daa
networks_truth      <- readRDS("../data/networks_advice_las.rds")  # readRDS("../data/networks_truth.rds")[["3"]]$advice
networks_advice_css <- readRDS("../data/networks_advice_css.rds")
networks_sizes      <- readr::read_csv("../data-raw/Study1_Group sizes.csv")
```

```{r cssdata1, echo=FALSE, cache=TRUE}
# Calculating distances/similarity of true advice network vs CSS
css_rangeL <- lapply(names(networks_advice_css), function(n) {
  
  similR::similarity(
    c(list(networks_truth[[n]]), networks_advice_css[[n]]),
    statistic = statistics,
    normalized = FALSE, firstonly=TRUE, exclude_j=TRUE
  )
  
})

# Computing range
css_range <- lapply(css_rangeL, function(d) {
  
  d[, statistics, drop=FALSE] %>%
    as.data.frame %>%
    gather("statistic", "value") %>%
    group_by(statistic) %>%
    summarize(range = diff(range(value))) %>% 
    ungroup %>%
    spread(statistic, range)
  
}) %>%
  bind_rows %>%
  bind_cols(networks_sizes) %>%
  gather("statistic", "value", -Group, -groupSize) %>%
  group_by(statistic) %>%
  mutate(
    groupSize = as.factor(groupSize),
    value_01  = (value - min(value))/diff(range(value))
  ) %>% ungroup

# Computing Max
css_max <- lapply(seq_along(css_rangeL), function(d) {
  
  css_rangeL[[d]][, statistics, drop=FALSE] %>%
    as.data.frame %>%
    gather("statistic", "value") %>%
    group_by(statistic) %>%
    summarize(
      max = max(value),
      min = min(value)
      ) %>% 
    mutate(
      group_id = networks_sizes[d,"Group",drop=TRUE]
    ) %>%
    ungroup # %>%
    #spread(statistic, max)
  
}) %>%
  bind_rows

dcast(setDT(css_max), group_id~ statistic, value.var=c("max", "min"))


```


```{r violin-plot, fig.cap="Distribution of Within Group Ranges of Similarity (S) and Distance (D) Statistics with respect to the LAS. Values are normalized to range between 0 and 1.", cache=TRUE}
library(ggplot2)
css_range %>%
  
  # Adding nicer labels
  mutate(statistic = names(statistics)[match(statistic, statistics)]) %>%
  
  # Plot
  ggplot(aes(x = statistic, y = value_01)) +
  geom_violin() +
  geom_jitter(height = 0, width=.1, aes(colour=groupSize, shape = groupSize), size=4) +
  scale_colour_viridis_d(alpha = .7)  +
  
  labs(y = "Normalized statistic", x = "", shape = "Group Size", colour="Group Size") +
  theme(axis.text = element_text(angle = 45, hjust = 1))
```


```{r corrplot, cache=TRUE}
# Calcularing correlation
css_corr <- css_range %>%
  select(-value_01) %>%
  spread(statistic, value) %>%
  select(-Group, -groupSize) %>%
  cor

# Adding nicer labels
css_corr <- names(statistics)[match(rownames(css_corr), statistics)] %>%
  replicate(2, ., simplify = FALSE) %>%
  `dimnames<-`(css_corr, .)

# Gathering the data
css_corr <- css_corr %>%
  as_tibble %>%
  cbind(a = rownames(css_corr), .) %>%
  gather("b", "value", -a)

# Plotting
ggplot(css_corr, aes(x=a, y=b, fill=value))+
  scale_fill_distiller(palette = "RdBu") +
  geom_bin2d() +
  labs(fill     = "Correlation") +
  labs(title    = "Correlation between similarity statistics") +
  theme(
    axis.text       = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    axis.title      = element_blank()
    )
```

Observation: The hamming normalized seems to be the same as the Mean manhattan (s)

## Within groups variability

```{r cssdata-within, echo=FALSE, cache=TRUE}
# Filling the gaps
groups_ids <- names(networks_advice_css) %>%
  unique

networks_advice_css_filled <- networks_advice_css
for (g in groups_ids) {
  
  # Adding i's perception from the true network data
  for (i in names(networks_advice_css[[g]])) 
    networks_advice_css_filled[[g]][[i]][i,] <- networks_truth[[g]][i,]
  
}

# Calculating distances/similarity of true advice network vs CSS
css_range_within <- lapply(names(networks_advice_css), function(n) {
  
  similR::similarity(
    networks_advice_css_filled[[n]],
    statistic = statistics,
    normalized = FALSE, firstonly=FALSE, exclude_j=FALSE
  )
  
})

# Computing range
css_range_within <- lapply(css_range_within, function(d) {
  
  d[, statistics, drop=FALSE] %>%
    as.data.frame %>%
    gather("statistic", "value") %>%
    group_by(statistic) %>%
    summarize(range = diff(range(value))) %>% 
    ungroup %>%
    spread(statistic, range)
  
}) %>%
  bind_rows %>%
  bind_cols(networks_sizes) %>%
  gather("statistic", "value", -Group, -groupSize) %>%
  group_by(statistic) %>%
  mutate(
    groupSize = as.factor(groupSize),
    value_01  = (value - min(value))/diff(range(value))
  ) %>% ungroup


```


```{r violin-plot-within, fig.cap="Distribution of Within Group Ranges of Similarity (S) and Distance (D) Statistics within the groups. Values are normalized to range between 0 and 1.", cache=TRUE}
css_range_within %>%
  
  # Adding nicer labels
  mutate(statistic = names(statistics)[match(statistic, statistics)]) %>%
  
  # Plot
  ggplot(aes(x = statistic, y = value_01)) +
  geom_violin() +
  geom_jitter(height = 0, width=.1, aes(colour=groupSize, shape = groupSize), size=4) +
  scale_colour_viridis_d(alpha = .7)  +
  
  labs(y = "Normalized statistic", x = "", shape = "Group Size", colour="Group Size") +
  theme(axis.text = element_text(angle = 45, hjust = 1))
```


# TCI analysis

```{r model-data}
# List of variables to run regressions
tci_variables <- c("CI_avg.T1", "CI_avg.T2")
covariates    <- c("GrpProportionMale", "GrpRangeAge", "GrpProportionNonwhite",
                   "GrpSize", "GPA_AVG")
```


## LAS

```{r model-data-las}
# Regrouping the css_range
model_data <- css_range %>%
  ungroup %>%
  select(-value_01) %>%
  spread(statistic, value) %>%
  select(-groupSize)

model_data <- dat_group %>%
  select(Group, !!tci_variables, !!covariates) %>%
  mutate(Group = as.integer(Group)) %>%
  left_join(model_data, by = "Group")
```


Descriptive statistics

```{r, fig.cap="Descriptive Statistics "}
model_data %>%
  select(GrpSize, !!statistics) %>%
  gather(Statistic, Value, -GrpSize) %>%
  group_by(Statistic, GrpSize) %>%
  summarise(
    N       = n(),
    Min      = min(Value),
    Max      = max(Value),
    Variance = var(Value)
  ) %>%
  ungroup %>%
  arrange(Statistic, GrpSize) %>%
  knitr::kable(., digits = 2L)
```

```{r, fig.cap="Descriptive Statistics2 "}
model_data %>%
  select(GrpSize, !!statistics) %>%
  gather(Statistic, Value, -GrpSize) %>%
  group_by(Statistic, GrpSize) %>%
  mutate(
    Type1 = quantile(Value, 1/3),
    Type2 = quantile(Value, 2/3),
    Type  = (Value < Type1) + (Value < Type2)
    ) %>%
  ungroup %>%
  group_by(Type, Statistic) %>%
  summarise(
    N       = n(),
    Min      = min(Value),
    Max      = max(Value),
    Variance = var(Value)
  ) # %>%
  arrange(Statistic, GrpSize) %>%
  knitr::kable(., digits = 2L)
```


```{r full-model}
mnum <- 0L
models <- vector("list", 0)
for (s in statistics) {
  
  m <- paste("`CI_avg.T2` ~ GrpProportionMale + GrpRangeAge + GrpProportionNonwhite + 
             GPA_AVG + factor(GrpSize)+",s)
  models[[mnum <- mnum + 1]] <- lm(as.formula(m), data = model_data)
  
}
```

```{r full-model-less03}

model_data <- filter(model_data, GrpSize != 3L) %>%
  rename(`CI_avg.T2(-3)` = CI_avg.T2)

for (s in statistics) {
  
  m <- paste("`CI_avg.T2(-3)` ~ GrpProportionMale + GrpRangeAge + GrpProportionNonwhite + 
             GPA_AVG + factor(GrpSize)+",s)
  models[[mnum <- mnum + 1]] <- lm(as.formula(m), data = model_data)
  
}
```

\clearpage 

\blandscape

```{r table, results='asis'}
# Generating the table
tabfun <- if (knitr::is_html_output()) {
  texreg::htmlreg
  } else {
    function(...) texreg::texreg(...)
  }

# How many variables that are not statistics
ncovars <- length(setdiff(names(models[[1]]$coefficients), statistics)) + 1L

tab <- capture.output(tabfun(
  models,
  # dep.var.labels   = c("All groups", "Groups of size 4 and 5"),
  # dep.var.caption  = "Standarized TCI",
  custom.coef.map = c(list(
    "GrpProportionMale"= "Prop. of Males" ,
    "GrpProportionNonwhite" = "Prop. of Non-white",
    "GrpRangeAge"      = "Age Range" ,
    "GPA_AVG"          = "Avg GPA",
    "factor(GrpSize)4" = "Size = 4" , 
    "factor(GrpSize)5" = "Size = 5" ,
    "(Intercept)"      = "(Intercept)"
    ), structure(names(statistics), names = statistics)
    ),
  groups  = list(
    "Distance/Similarity" = ncovars:(ncovars + length(statistics) - 1)),
  caption = "Regression using different distance/similarity measurements"
  ))

cat(tab, sep="\n")
```

\elandscape

\clearpage 

\normalsize


## Within


```{r model-data-within}
# Regrouping the css_range
model_data <- css_range_within %>%
  ungroup %>%
  select(-value_01) %>%
  spread(statistic, value) %>%
  select(-groupSize)

model_data <- dat_group %>%
  select(Group, !!tci_variables, !!covariates) %>%
  mutate(Group = as.integer(Group)) %>%
  left_join(model_data, by = "Group")
```

```{r full-model-within}
mnum <- 0L
models <- vector("list", 0)
for (s in statistics) {
  
  m <- paste("`CI_avg.T2` ~ GrpProportionMale + GrpRangeAge + GrpProportionNonwhite + 
             GPA_AVG + factor(GrpSize)+",s)
  models[[mnum <- mnum + 1]] <- lm(as.formula(m), data = model_data)
  
}
```

```{r full-model-less03-within}

model_data <- filter(model_data, GrpSize != 3L) %>%
  rename(`CI_avg.T2(-3)` = CI_avg.T2)

for (s in statistics) {
  
  m <- paste("`CI_avg.T2(-3)` ~ GrpProportionMale + GrpRangeAge + GrpProportionNonwhite + 
             GPA_AVG + factor(GrpSize)+",s)
  models[[mnum <- mnum + 1]] <- lm(as.formula(m), data = model_data)
  
}
```

\clearpage 

\blandscape

```{r table-within, results='asis'}
# Generating the table
tabfun <- if (knitr::is_html_output()) {
  texreg::htmlreg
  } else {
    function(...) texreg::texreg(...)
  }

# How many variables that are not statistics
ncovars <- length(setdiff(names(models[[1]]$coefficients), statistics)) + 1L

tab <- capture.output(tabfun(
  models,
  # dep.var.labels   = c("All groups", "Groups of size 4 and 5"),
  # dep.var.caption  = "Standarized TCI",
  custom.coef.map = c(list(
    "GrpProportionMale"= "Prop. of Males" ,
    "GrpProportionNonwhite" = "Prop. of Non-white",
    "GrpRangeAge"      = "Age Range" ,
    "GPA_AVG"          = "Avg GPA",
    "factor(GrpSize)4" = "Size = 4" , 
    "factor(GrpSize)5" = "Size = 5" ,
    "(Intercept)"      = "(Intercept)"
    ), structure(names(statistics), names = statistics)
    ),
  groups  = list(
    "Distance/Similarity" = ncovars:(ncovars + length(statistics) - 1)),
  caption = "Regression using different distance/similarity measurements"
  ))

cat(tab, sep="\n")
```

\elandscape

\clearpage 

\normalsize


# Network plots


```{r plot-graph-3, fig.cap="LAS Networks of size 3", echo=FALSE, results='hide'}
networks_size3 <- networks_truth[networks_sizes$groupSize==3] %>%
  lapply(igraph::graph_from_adjacency_matrix)

nnetworks_size3 <- length(networks_size3)
op <- par(mfrow=c(3,3), mai=rep(.2, 4))
lapply(networks_size3, plot, vertex.size=10, vertex.label="")
par(op)
```

```{r plot-graph-4, fig.cap="LAS Networks of size 4.", echo=FALSE, results='hide'}
networks_size4 <- networks_truth[networks_sizes$groupSize==4] %>%
  lapply(., igraph::graph_from_adjacency_matrix)

nnetworks_size4 <- length(networks_size4)
op <- par(mfrow=c(4,5), mai=rep(.2, 4))
lapply(networks_size4, plot, vertex.size=10, vertex.label="")
par(op)
```


```{r plot-graph-5, fig.cap="LAS Networks of size 5.", echo=FALSE, results='hide'}
networks_size5 <- networks_truth[networks_sizes$groupSize==5] %>%
  lapply(., igraph::graph_from_adjacency_matrix)

nnetworks_size5 <- length(networks_size5)
op <- par(mfrow=c(4,5), mai=rep(.2, 4))
lapply(networks_size5, plot, vertex.size=10, vertex.label="")
par(op)
```

<!-- # Factor analysis -->

<!-- ```{r analysis} -->
<!-- dat_individual %>% -->
<!--   select( -->
<!--     PID, Group, -->
<!--     RMEscore, -->
<!--     SI3Fac1, SI3Fac2, SI3Fac3, -->
<!--     # ZRMEscore, ZFLAbRel, -->
<!--     FL_absolute_avg -->
<!--     ) -->
<!-- ``` -->




